Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2013-01-07 chi2013-webbox.tex
mum edits

Second, we conducted a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these types of sources.  We examined data sources in six common domains: contacts, events, music, shopping, social networking and weather. Our analysis found that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibit among more complex schemas, such as online retailers' product catalogues.

Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.  It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.  In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.  We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.

@@

The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.  For example, the uniform query capabilities provided by SEMEX~\cite{semex} and \emph{Stuff I've Seen}\cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.  Similarly, Haystack \cite{haystack} demonstrated the reduced effort and increased efficiency of being able to view and re-use information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.

@@

Our approach is most similar to systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but that avoid the need for any sort of ``programming'' by the user.  Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.  Our system, \emph{DataPalette} extends this approach to a data workbench environment which aims to allow people to easily consolidate and compare instances by their properties.

@@

\begin{table*}[htb]
\small
\begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
				&				& {\bf average}				& {\bf average }				& 				& {\bf prop name}			& {\bf structural} 			& {\bf model } \\
				& {\bf source}		& {\bf props/record}			& {\bf depth/record}			& {\bf equivalent to}	& {\bf inconsistent}			& {\bf inconsistency} 			& {\bf inconsistency} \\
\hline
social networking	& google+			& 12						& 1.52					& tw:5, li:4, fb:4		& tw:3, li:2, fb:2				& tw:0, li:1, fb:1				& tw:0, li:0, fb:0 \\
(profiles)			& twitter			& 39						& 1.37					& li:4, fb:4 			& li:2, fb:1					& li:0, fb:0					& li:0, fb:0 \\
				& linkedin			& 18						& 2.42					& fb:7			& fb:5					& fb:3					& fb:0 \\
				& facebook		& 22						& 1.5						&				&						&						& \\
\hline
calendars			& google calendar	& 15						& 1.46					& eb:9			& eb:6					& eb:1					& eb:1 \\
(events)			& eventbrite		& 32						& 1.43					&				&						& 						& \\
\hline
retail				& milo			& 5						& 1.2						& z:2, s:3, e:1, a:3	& z:2, s:2, e:1, a:3			& z:0, s:0, e:0, a:2			& z:0, s:0, e:0, a:0 \\
(products)			& zappos			& 6						& 1						& s:2, e:2, a:3		& s:2, e:2, a:2				& s:0, e:0, a:0				& s:0, e:0, a:0 \\
				& shopping.com	& 11						& 3.74					& e:2, a:8			& e:2, a:8					& e:0, a:0					& e:0, a:0 \\
				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
				& amazon			& 108					& 1.96					&				& 						&						& \\
\hline
contacts			& google contacts	& 7						& 1.46					& yp:2			& yp:2					& yp: 0					& yp: 0 \\
				& yellowpages		& 25						& 1						&				&						&						& \\
\hline
music			& echonest		& 4						& 1.83					& so:3, r:3, sp:3		& so:3, r:3, sp:3				& so:0, r:0, sp:0				& so:0, r:0, sp:0 \\
(songs/			& soundcloud		& 42						& 2.14					& r:12, sp:7		& r:11, sp:6 				& r:1, sp:1					& r:1, sp:0 \\
tracks)			& rdio			& 27						& 2						& sp: 7			& sp: 6					& sp: 1					& sp: 1 \\
				& spotify			& 10						& 2.44					&				&						&						& \\
\hline
weather			& weatherbug		& 8						& 2						& wu: 5, y: 5		& wu: 5, y: 5				& wu: 5, y: 4				& wu: 0, y: 0 \\
(forecasts)		& wundergound	& 52						& 1.38					& y:12			& y:12					& y:7						& y:3 \\
				& yahoo weather	& 22						& 1.58					&				&						&						& \\
\end{tabular}
\caption{\emph{PS.2 Analysis of Data Feeds}: Structural statistics about data sampled from each API organised by category.  The first column lists the category and type of data record that was examined.  Columns 3 and 4 list the average number of properties per data record, and number of levels from root to leaf. The remaining fields show results from a pairwise analysis of data records from each of the data sources with those of the others in each category; values are prefixed with the first letter(s) of the target schema the row schema was compared against. For example ``Tw:5'' in the 5th column of Google+ row indicates that Facebook and Twitter records had 5 equivalent fields. }\label{tbl:prestudy2}
\end{table*}

@@

\subsubsection{Results}
Table \ref{tbl:prestudy2} shows the raw results of the process described.  A number of features were quite striking about the data; First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest. In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories. Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.  While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.  Only Spotify, out of the 4 music providers, included information about the album featuring a given song.  Such omissions show us that these data APIs reflect each provider's own specific needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.

>>>>

Second, we carried out a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these multiple sources.  For this, we examined data sources from six common domains: contacts, events, music, shopping, social networking and weather. We discovered in our analysis that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibited among more complex schemas, such as online retailers' product catalogues.

Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' eliminating the need for people to write bespoke code to effectively combine and compare heterogeneous data.  This interface was principally designed to support people's data integration needs for data heterogeneity, using observations made during the pre-studies. In particular, DataPalette was designed to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.  We then conducted a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we had devised, and that it effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.

@@

The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.  For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.  Similarly, Haystack \cite{haystack} demonstrated the benefit of viewing and re-using information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.

@@

Our approach is most similar to systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but that avoid the need for any sort of ``programming'' by the user.  Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.  Our system, \emph{DataPalette} extends this approach to a data workbench environment which allows users to easily consolidate and compare instances by their properties.

@@

@@

\subsubsection{Results}
While most of the data records were small and relatively simple, nearly one in each category provided access to substantially larger and more complex records than the rest. In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories. Examining equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.  While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.  Only Spotify, out of the 4 music providers, included information about the album featuring a given song.  Such omissions show us that these data APIs reflect each provider's own specific needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
